JD notes on effects

### 2022-06-16 manuscript/body.tex
Autocommit

\begin{figure}[h]

@@

\begin{figure}[h]
\begin{center}
\includegraphics{justify_preds.isolate.pdf}
\end{center}
\caption{{\bf A comparison of conventional $95\%$ confidence bands and effects.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence bands around the predcitions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal input variable, i.e., the model center, corresponds to the effects of the focal input variable from \pkg{varpred}. For simple models, the isolated confidence bands crosses at the model center. The wider curves incorporates not only the uncertainties due to intercept term but also other non-focal input variables. On the other hand, effects (narrower crossing lines) only take into account the main effect uncertainty of the focal input variable.}
\label{fig:justify_ci_plots}
\end{figure}

@@

\begin{figure}[h]
\begin{center}
\includegraphics{justify_anchors.combined.pdf}
\end{center}
\caption{{\bf A comparison of min-based and center- (mean-) based anchors.} The center-based effects are based on the mean of the focal predictor while the min-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines).} 
\label{fig:justify_anchors}
\end{figure}

@@

\subsection{Predictions-styled and effects-styled plots}

As mentioned above, in some applications, one may be interested in uncertainties associated with the focal predictor only. In that case, it is important to exclude all the uncertainties as a result of non-focal predictors and only show confidence intervals describing the predictor of interest. To illustrate this, we consider the simulation model described in \nameref{S1_Appendix}.

We start by generating two predictions from the fitted model \nameref{S1_Appendix}: 1) one with cubic polynomial of \code{age} as a focal predictor; and 2) one with a simple focal predictor (non-polynomial) of \code{Wealthindex}. Then we compared non-anchored and mean-anchored (effects) confidence intervals around the predicted values (see Fig~\ref{fig:pred_cubic_plots}).

\begin{figure}[h]
\begin{center}
\includegraphics{cubic_predictors_preds.ggp.pdf}
\end{center}
\caption{{\bf Prediction-styled and effect-styled plots for cubic polynomial and simple focal predictors.} A: The focal predictor is a cubic polynomial. B: The focal predictor is a linear function, with a cubic polynomial non-focal predictor. In both cases, the mean-anchored effects are narrower than the non-anchored (predictions), but intersects at the mean of the focal predictor (model center) only in the case of linear focal predictor -- B. In the case of cubic polynomial focal predictor, the model center is not a point in the predictor space, i.e., mean of the focal predictor, but a combination of all the model variables in the cubic polynomial, consequently, mean-anchored curves will not necessarily cross at a point -- A. The horizontal black and red dashed lines are the observed and predicted average household size, i.e., $\over{hh~size}$ and $\over{\widehat{hh~size}}$, respectively; the inner red dotted curves and crossing lines are the effects for A and B, respectively; while the outer blue curves are the \BC{predictions (place holder for now)} in both cases. The central curve and trend line, in A and B, respectively; are the predicted household size based on age and wealth index, respectively. \bmb{Use different line types ...}\BC{Confirm with BMB}}
\label{fig:pred_cubic_plots}
\end{figure}

@@

\begin{figure}[h]
\begin{center}
\includegraphics{glm_two_predictor_preds.ggp.pdf}
\end{center}
\caption{{\bf Mean-based and whole-sample-based predictions.} The two approaches produce relatively different estimates in each of the predictors but the difference is more pronounced when the effect of the particular focal predictor has no strong effect -- \emph{on the left}. In both cases, \emph{right} and \emph{left}, the average predicted probability of improved water based on whole-sample-based approach is very close to the observed proportion in the simulated data. The horizontal blue, black and the red dotted lines are the average mean-based, whole-sample-based predicted probability and observed proportion of improved water quality, respectively. The vertical dotted black line represents the mean of the focal predictor, and the point at which it crosses the red dotted line represents the expected ``perfect'' prediction at the model center. The grey points are binned observations -- observed proportions of improved water quality in each bin. We would expect all the curves, horizontal, and vertical lines to cross at the model center but due the nonlinear averaging (at least in the bias corrected whole-sample-based approach) of the predictions on the response scale, this may not always be the case.} 
\label{fig:pred_bin_plots}
\end{figure}

@@

\begin{figure}[h]
\begin{center}
\includegraphics{mediate_preds.ggp.pdf}
\end{center}
\caption{{\bf Mean-based and whole-sample-based predictions for mediated effects.} These figures compare the mean-based and whole-sample-based predictions for the mediated (multivariate) and non-mediated (univariate) models. A: In the absence of mediator variable, both mean-based and whole-sample-based approaches predictors are identical since there is no any other additional (non-focal predictors) sources of bias. Consequently, for both approaches, the average predicted probability is very close to the observed proportion. B: When the mediator variable is included, there is no direct effect of $x$ on $z$ and as a result the predictions do not necessarily align with the observations. However, the whole-sample-based approach still closely approximates the average proportion in the simulated data. The horizontal blue, black and the red dotted lines are the average mean-based, whole-sample-based predicted probability and observed proportion, respectively. The vertical dotted black line represents the mean of the focal predictor, and the point at which it crosses the red dotted line represents the expected ``perfect'' prediction at the model center. The grey points are the binned observations.}
\label{fig:pred_mediated_plots}
\end{figure}

>>>>

\begin{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{justify_preds.isolate.pdf}
\end{center}
\caption{{\bf A comparison of conventional $95\%$ confidence bands and effects.} The description of horizontal, vertical and trend lines remain the same as above. The wider dotted blue curves overlaying the green curves corresponds to the conventional confidence bands around the predcitions from \pkg{emmeans} and \pkg{effects}, while the narrower black curves crossing at the mean of the focal input variable, i.e., the model center, corresponds to the effects of the focal input variable from \pkg{varpred}. For simple models, the isolated confidence bands crosses at the model center. The wider curves incorporates not only the uncertainties due to intercept term but also other non-focal input variables. On the other hand, effects (narrower crossing lines) only take into account the main effect uncertainty of the focal input variable.}
\label{fig:justify_ci_plots}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{justify_anchors.combined.pdf}
\end{center}
\caption{{\bf A comparison of min-based and center- (mean-) based anchors.} The center-based effects are based on the mean of the focal predictor while the min-anchored effects are based on the minimum of the focal predictor. As we mentioned earlier, the choice of the anchor do not affect the main estimates (central trend lines).} 
\label{fig:justify_anchors}
\end{figure}

@@

\subsection{Prediction and effect styled plots}

In multivariate models, we may want to visualize the effect of a particular focal input variable. In this case, effects styled plots are preferable over the conventional prediction plots which generally focus on the \emph{total} effects. To illustrate this, we simulate a multivariate model described in \nameref{S1_Appendix}. We also use the same model to illustrate how predictions and effects differ in the presence of higher order interactions in focal or non-focal input variables.

We generated two sets of predictions and effects: 1) one with \code{age} as the focal input variable (a cubic polynomial polynomial focal predictor); and 2) one with \code{Wealthindex} as the focal in put variable (linear focal input variable). Then compared the predictions and effects as shown in Fig~\ref{fig:pred_cubic_plots}.

\begin{figure}
\begin{center}
\includegraphics{cubic_predictors_preds.ggp.pdf}
\end{center}
\caption{{\bf Prediction-styled and effect-styled plots for cubic polynomial and simple focal predictors.} A: The focal predictor is a cubic polynomial. B: The focal predictor is a linear function, with a cubic polynomial non-focal predictor. In both cases, the mean-anchored effects are narrower than the non-anchored (predictions), but intersects at the mean of the focal predictor (model center) only in the case of linear focal predictor -- B. In the case of cubic polynomial focal predictor, the model center is not a point in the predictor space, i.e., mean of the focal predictor, but a combination of all the model variables in the cubic polynomial, consequently, mean-anchored curves will not necessarily cross at a point -- A. The horizontal black and red dashed lines are the observed and predicted average household size, i.e., $\over{hh~size}$ and $\over{\widehat{hh~size}}$, respectively; the inner red dotted curves and crossing lines are the effects for A and B, respectively; while the outer blue curves are the \BC{predictions (place holder for now)} in both cases. The central curve and trend line, in A and B, respectively; are the predicted household size based on age and wealth index, respectively. \bmb{Use different line types ...}\BC{Confirm with BMB}}
\label{fig:pred_cubic_plots}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{glm_two_predictor_preds.ggp.pdf}
\end{center}
\caption{{\bf Mean-based and whole-sample-based predictions.} The two approaches produce relatively different estimates in each of the predictors but the difference is more pronounced when the effect of the particular focal predictor has no strong effect -- \emph{on the left}. In both cases, \emph{right} and \emph{left}, the average predicted probability of improved water based on whole-sample-based approach is very close to the observed proportion in the simulated data. The horizontal blue, black and the red dotted lines are the average mean-based, whole-sample-based predicted probability and observed proportion of improved water quality, respectively. The vertical dotted black line represents the mean of the focal predictor, and the point at which it crosses the red dotted line represents the expected ``perfect'' prediction at the model center. The grey points are binned observations -- observed proportions of improved water quality in each bin. We would expect all the curves, horizontal, and vertical lines to cross at the model center but due the nonlinear averaging (at least in the bias corrected whole-sample-based approach) of the predictions on the response scale, this may not always be the case.} 
\label{fig:pred_bin_plots}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics{mediate_preds.ggp.pdf}
\end{center}
\caption{{\bf Mean-based and whole-sample-based predictions for mediated effects.} These figures compare the mean-based and whole-sample-based predictions for the mediated (multivariate) and non-mediated (univariate) models. A: In the absence of mediator variable, both mean-based and whole-sample-based approaches predictors are identical since there is no any other additional (non-focal predictors) sources of bias. Consequently, for both approaches, the average predicted probability is very close to the observed proportion. B: When the mediator variable is included, there is no direct effect of $x$ on $z$ and as a result the predictions do not necessarily align with the observations. However, the whole-sample-based approach still closely approximates the average proportion in the simulated data. The horizontal blue, black and the red dotted lines are the average mean-based, whole-sample-based predicted probability and observed proportion, respectively. The vertical dotted black line represents the mean of the focal predictor, and the point at which it crosses the red dotted line represents the expected ``perfect'' prediction at the model center. The grey points are the binned observations.}
\label{fig:pred_mediated_plots}
\end{figure}
